Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-21 docs/proposals/NSF/ProjDesc/ProjDesc.tex
NM

Equation \eqref{eqn:StokesEq}, and its equivalent integral formulation, are microscopic descriptions required to model the shape evolution of individual grains due to fine-scale erosive effects, either mechanical or chemical. This microscopic description, however, becomes impractical when attempting to model seepage in very large porous systems where the microscopic geometry can be exceedingly complex. Instead, one must homogenize over microscopic details to arrive at an approximate description. Leading-order homogenization produces the well-known Darcy system 
\begin{align}
  \label{DarcyEq}
  \qq = - \bvec{K} \grad \phi
\end{align}
where $\qq$ is the {\em specific discharge} or {\em Darcy velocity}, and $\phi = z + p/(\rho g)$ is the hydraulic head, with $\rho$ and $g$ representing the fluid density and gravitational constant respectively.  Above, $\bvec{K}$ is the {\em conductivity}, which is related to the {\em permeability} $\bvec{k}$ through $\bvec{K} = \bvec{k} \rho g/\mu$.  Both the conductivity and permeability are in general tensors, and they reduce to scalars in the case of an isotropic medium. Note that the Darcy velocity is related to the tracer velocity through $\qq = \theta_f \uu$, where $\theta_f$ is porosity (i.e.~fluid volume fraction).  In the case of an isotropic medium, the Kozeny-Carman equation is often used to relate permeability to volume fraction~\cite{bear2013dynamics}
% See Bear Ch. 2.8, pg. 111
\begin{align}
  k = C \frac{\theta_f^3}{(1-\theta_f)^2}.
\end{align}
The macroscopic description offered by Eq.~\eqref{DarcyEq}, although approximate, simplifies analysis tremendously in that individual grains are no longer resolved and thus the domain geometry is generally much simpler.

@@

More specifically, we plan to run large batches of the high-fidelity boundary-integral-equation-based numerical simulations of erosion and then leverage rapid advances in the field of machine learning and deep-neural networks \cite{ling2016reynolds, kutz2017deep, ling2016machine, pelt2018mixed} to develop reduced models that parametrize the effects of erosion on the evolution of macroscopic medium properties and the flow.  The output of the high-fidelity simulations of the grain configuration can be transformed into essentially an image format, permitting use of cutting-edge advances in deep neural-networks for image analysis~\cite{pelt2018mixed}. Some important advantages of referenced work are the use of dilated convolutions in a deeply connected network, which first mixes scales and second reduces the number of super-parameters present, thus largely mitigating the risk of overfitting data. Additionally, high-fidelity flow conditions can be used to train a neural network---a strategy that has been successfully used by others to predict, for example, an anisotropy tensor~\cite{ling2016reynolds}.  Both PIs will supervise a graduate student who will design and implement deep-neural networks for learning the essence of the high-fidelity simulations. Once the essential features are learned, they will be incorporated into reduced-order models that harness the multiphase framework. These reduced models will enable more efficient and larger scale coarse-grained simulations of dynamic erosion.

%----------------------------------------------------------------------%
\subsection{Experimental Validation}
\label{sec:experiments}
As mentioned previously, researchers at FSU's Geophysical Fluid Dynamics
Institute (GFDI) have investigated the mechanisms behind sinkhole
formation with a series of scaled-down laboratory
experiments~\cite{tao2014experimental}. First, already existing data
from these experiments offers a testbed for comparison against our newly
developed simulations. Furthermore, {\bf we propose to run a set of new
experiments at the facilities offered by GFDI}, spearheaded by an undergraduate student in close collaboration with the graduate students, PIs, and an experimentalist at GFDI. These new experiments will more deeply probe the mechanisms underlying sinkhole formation with the special design goal of comparison against the numerical simulations. These experiments will be used to calibrate and validate our computations.
\todo[inline]{We have more room at the end of the document. This
section seems a bit brief and could maybe be expanded.}

>>>>

Equation \eqref{eqn:StokesEq}, and its equivalent integral formulation, are microscopic descriptions required to model the shape evolution of individual fragments and grains due to fine-scale erosive effects, either mechanical or chemical. This microscopic description, however, becomes impractical when attempting to model seepage in very large porous systems where the microscopic geometry can be exceedingly complex. Instead, one must homogenize over microscopic details to arrive at an approximate description. Leading-order homogenization produces the well-known Darcy system 
\begin{align}
  \label{DarcyEq}
  \qq = - \bvec{K} \grad \phi
\end{align}
where $\qq$ is the {\em specific discharge} or {\em Darcy velocity}, and $\phi = z + p/(\rho g)$ is the hydraulic head, with $\rho$ and $g$ representing the fluid density and gravitational constant respectively.  Above, $\bvec{K}$ is the {\em conductivity}, which is related to the {\em permeability} $\bvec{k}$ through $\bvec{K} = \bvec{k} \rho g/\mu$.  Both the conductivity and permeability are in general tensors, and they reduce to scalars in the case of an isotropic medium. Note that the Darcy velocity is related to the tracer velocity through $\qq = \theta_f \uu$, where $\theta_f$ is porosity (i.e.~fluid volume fraction).  In the case of an isotropic medium, the Kozeny-Carman equation is often used to relate permeability to volume fraction~\cite{bear2013dynamics}
% See Bear Ch. 2.8, pg. 111
\begin{align}
  k = C \frac{\theta_f^3}{(1-\theta_f)^2}.
\end{align}
The macroscopic description offered by Eq.~\eqref{DarcyEq}, although approximate, simplifies analysis tremendously in that individual grains are no longer resolved and thus the domain geometry is generally much simpler.

@@

{\bf More specifically, we plan to run large batches of the high-fidelity boundary-integral-equation-based numerical simulations of erosion and then leverage rapid advances in the field of machine learning and deep-neural networks \cite{ling2016reynolds, kutz2017deep, ling2016machine, pelt2018mixed} to develop reduced models that parametrize the effects of erosion on the evolution of macroscopic medium properties and the flow.}  The output of the high-fidelity simulations of the grain configuration can be transformed into essentially an image format, permitting use of cutting-edge advances in deep neural-networks for image analysis~\cite{pelt2018mixed}. Some important advantages of referenced work are the use of dilated convolutions in a deeply connected network, which first mixes scales and second reduces the number of super-parameters present, thus largely mitigating the risk of overfitting data. Additionally, high-fidelity flow conditions can be used to train a neural network---a strategy that has been successfully used by others to predict, for example, an anisotropy tensor~\cite{ling2016reynolds}.  Both PIs will supervise a graduate student who will design and implement deep-neural networks for learning the essence of the high-fidelity simulations. Once the essential features are learned, they will be incorporated into reduced-order models that harness the multiphase framework. These reduced models will enable more efficient and larger scale simulations of dynamic erosion than are feasible with the high-fidelity boundary-integral simulations.

%----------------------------------------------------------------------%
\subsection{Experimental Validation}
\label{sec:experiments}

	As mentioned previously, researchers at FSU's Geophysical Fluid Dynamics Institute (GFDI) have investigated the mechanisms behind sinkhole formation with a series of scaled-down laboratory experiments~\cite{tao2014experimental}. We plan to make use of the knowledge gained from these past experiments, as well as the facilities and equipment already in place, to inform and validate our computational program.

	First, already existing data from these experiments offers a wonderful testbed for comparison against the newly developed simulations. This data provides information about the physical conditions and broad parameter regimes required for sinkhole collapse. Second, and more importantly, we propose to conduct a new campaign of laboratory experiments at the GFDI facilities which will be spearheaded by PI Moore and an undergraduate student funded by the project. These experiments will make use of much of the same infrastructure, as well as technical staff present at GFDI with working knowledge of the specific equipment. The experiments will be conducted in close cooperation with the graduate students to ensure constant feedback between advances in the computational program and experiments.

	The proposed experiments involve two parallel lines of inquiry that we plan to merge in later stages of the project. The first line of inquiry involves the dynamic erosion of a porous material by either mechanical or chemical action. In this context, the porous medium will be represented by fragments of either erodible material, such as bentonite clay, or dissolve material, such as sucrose solids. We will perform controlled experiments to examine how important macroscopic properties, such as permeability, mechanical strength, and anisotropy, evolve dynamically as the constituents disintegrate. These experiments will make use of Darcy-column containers that allow precise control of flux through the medium and precise hydraulic head measurements to infer permeability. These Darcy columns have already been fabricated and are available for use at GFDI.

	The second line of inquiry will extend the sinkhole experiments that were previously conducted at GFDI, with the specific aim of informing and validating our numerical simulations. These experiments will build on the significant infrastructure already in place from the previous sinkhole experiments, with the express consent of our collaborator Prof. Ming Ye (see collaboration letter). This infrastructure permits precise control of the hydraulic head in a confined and unconfined aquifers -- it is the difference in the heads creates conditions favorable for sinkhole formation. The new experiments will use both natural (sand) and artificial (glass beads) granular materials. The advantage of the latter is to achieve a mono-disperse medium for direct comparison against numerical simulations. Time permitting, we hope to merge the two lines of inquiry to examine the influence of material erodibility on gravitational reconfiguration and specifically sinkhole formation. The materials available for laboratory experiments, in particular the relatively rapid rates of degradation that can be achieved, would enable us to study phenomenon that occur on much longer timescales in the natural world.

\todo[inline]{Like it?}
